% ============================================================
% Section 4 — Cross-Suite Analysis
% Requires: \usepackage{booktabs}
% ============================================================

\section{Cross-Suite Analysis}\label{sec:analysis}

\subsection{Analysis Methodology}\label{sec:metrics}

We characterise each benchmark along four categories of properties.

\emph{Static rule-base properties} (Table~\ref{tab:static}) are extracted by parsing the DRL source text.
Conditions per rule are counted by matching LHS fact patterns; the alpha-sharing ratio is
$1 - (\mathrm{distinct\ fact\ types} / \mathrm{total\ pattern\ occurrences})$, quantifying how much
node sharing the RETE alpha network can exploit.
Temporal operators (\texttt{after}, \texttt{before}, \texttt{accumulate}, \texttt{window:time},
\texttt{not}, \texttt{eval}) are tallied by regex over the DRL text.
Maximum forward-chain depth is computed as the longest path on the SCC condensation DAG of the
rule dependency graph, equivalent to standard Datalog stratification~\cite{ullman1989}.

\emph{Dataset properties} (Table~\ref{tab:stream}) are computed from the event stream before replay.
The inter-arrival time~(IAT) coefficient of variation (CV) is $\sigma_{\mathrm{IAT}} / \mu_{\mathrm{IAT}}$,
computed per entity (aircraft, symbol, or user) and then aggregated; CV~$>$~1 classifies the
stream as \emph{bursty}.

\emph{Runtime dynamics and coverage-variant analysis} (Table~\ref{tab:runtime}) are measured by replaying the event dataset through a Drools stream-mode rule session whose internal clock is advanced to each event's original timestamp, preserving real-world inter-arrival patterns. This analysis is performed on the initial baseline dataset and two reduced-coverage dataset variants per domain, which are produced by modifying event field values to progressively suppress pipeline entry conditions while keeping the rule base and event count (${\sim}1.6$M) unchanged. We report peak working memory size (maximum facts held simultaneously), rules fired per event, throughput, latency, and rule coverage.

\emph{Forward-chain depth activations} (Table~\ref{tab:depth}) are measured for the initial baseline dataset only. We assign each rule its depth $d$ in the dependency graph and report $P(d)$, the fraction of total activations at each depth.

\begin{table}[t]
\centering
\caption{Static rule-base and structural properties.}\label{tab:static}
\small
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Property} & \textbf{AirTraffic} & \textbf{Binance} & \textbf{Wikimedia} \\
\midrule
Rules (R)                       & 103      & 108      & 70      \\
Avg / max conditions per rule   & 0.27 / 2 & 2.02 / 4 & 0.01 / 1 \\
Alpha sharing ratio             & 0.464    & 0.899    & 0.738   \\
Distinct input fact types       & 15       & 22       & 32      \\
Max forward-chain depth         & 9        & 4        & 5       \\
Temporal operators              & 15       & 35       & 44      \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\caption{Dataset and stream properties.}\label{tab:stream}
\small
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Property} & \textbf{AirTraffic} & \textbf{Binance} & \textbf{Wikimedia} \\
\midrule
Dataset size (events)         & 1,600,000   & 1,613,159   & 1,599,966   \\
Time span                     & 19.2~min    & 30.0~min    & 4.15~days   \\
Distinct entities             & 14,301 acft & 10 symbols  & 32,225 users \\
Arrival rate (ev/s)           & 1,391       & 896         & 4.5         \\
IAT coeff.\ of variation (CV) & 1.156       & 4.203       & 9.958       \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\caption{Forward-chain depth activation probability $P(d)$: fraction of depth-assigned
activations at each depth. Cross-cutting WM monitors (R098, R099) are excluded from the
AirTraffic distribution; Binance and Wikimedia terminate at depths~3 and~5
respectively.}\label{tab:depth}
\small
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}lcccccccc@{}}
\toprule
\textbf{Domain} & \textbf{d=0} & \textbf{d=1} & \textbf{d=2} & \textbf{d=3}
  & \textbf{d=4} & \textbf{d=5} & \textbf{d=6} & \textbf{d=7--9} \\
\midrule
AirTraffic & 16.9\% & 17.0\% & 12.7\% & 2.3\% & 0.2\% & 3.0\% & 47.7\% & 0.3\% \\
Binance    & 49.0\% & 38.8\% &  6.9\% & 5.3\% & ---   & ---   & ---    & ---   \\
Wikimedia  & 69.3\% & 29.8\% & 0.22\% & 0.22\%& 0.22\%& 0.22\%& ---    & ---   \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t]
\caption{Runtime dynamics and coverage-variant analysis. All variants preserve ${\sim}1.6$M events.}\label{tab:runtime}\label{sec:coverage-analysis}
\centering
\small
\setlength{\tabcolsep}{3.5pt}
\begin{tabular}{l rrr rrr rrr}
\toprule
 & \multicolumn{3}{c}{\textbf{AirTraffic}}
 & \multicolumn{3}{c}{\textbf{Binance}}
 & \multicolumn{3}{c}{\textbf{Wikimedia}} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
Metric & Base & Med & Low & Base & Med & Low & Base & Med & Low \\
\midrule
Coverage (\%)       & 71.8 & 68.9 & 28.2 & 63.9 & 56.5 & 44.4 & 78.6 & 57.1 & 24.3 \\
Rules fired\,/\,evt & 39.4 & 11.5 & 18.4 & 15.7 & 3.59 & 3.35 & 3.69 & 2.63 & 0.07 \\
Peak WM (facts)     & 31.7M & 12.0M & 18.7M & 172 & 172 & 162 & 13K & 12.3K & 6.3K \\
Throughput (ev/s)   & 2,636 & 7,127 & 13,147 & 23,515 & 23,821 & 23,518 & 1,891 & 44,297 & 76,812 \\
Latency (s)         & 607.0 & 224.5 & 121.7 & 68.0 & 67.2 & 68.0 & 845.9 & 36.1 & 20.8 \\
Peak Heap (MB)      & 16,286 & 2,872 & {---} & 73 & 73 & 73 & 73 & 73 & 73 \\
\bottomrule
\end{tabular}
\end{table}

Three observations emerge.
First, \textbf{coverage is driven by stream diversity, not volume}: in Binance, removing TRADE events (88.5\% of the dataset) reduces coverage by only 7.4\,pp, whereas the additional 1.1\% MARK events account for a further 12.1\,pp reduction.
Second, \textbf{cross-pipeline starvation cascades} are visible: in Wikimedia, disabling the Content pipeline eliminates \texttt{ArticleQuality} join partners required by five Vandalism-verification rules at depths~2--3, causing coverage loss beyond the directly disabled pipeline.
Third, \textbf{irreducible structural cores} define per-domain coverage floors: the Binance DEPTH backbone alone sustains 44.4\% coverage (48\,/\,108 rules), representing the minimum achievable without degrading the benchmark's architectural integrity.
