% ============================================================
% Related Work section — LNCS style
% Drop this into your main .tex between \section and the next \section.
% Requires: \usepackage{booktabs} for the table.
% BibTeX entries are in the companion related_work_section.md artifact.
% ============================================================

\section{Related Work}\label{sec:related}

\textbf{Rule-engine benchmarks} have historically focused on static reasoning over synthetic data.
The classic Waltz and Miss Manners programs~\cite{brownston1985ops5}, originally developed for OPS5~\cite{forgy1982rete}, remain standard micro-benchmarks for Rete-based engines including Drools, but test only pattern-matching throughput on fixed, synthetic workloads.
OpenRuleBench~\cite{liang2009openrulebench} evaluates production-rule, Datalog, and semantic-web engines across 11 workloads (LUBM, WordNet, DBLP), while RUBEN~\cite{angele2022ruben} updates and extends it as a reusable framework.
Bobek and Misiak~\cite{bobek2017benchmarking} generate synthetic rule models parameterised by rule count and condition length.
The Train Benchmark~\cite{szarnyas2018train} comes closest to continuous evaluation, testing incremental graph-pattern matching on synthetic railway models with Rete-derived engines---but targets model-transformation tools, not event streams.
None of these incorporate streaming data, temporal CEP operators, or structural rule-base analysis.
As Liang et al.\ note: \emph{``Rule engines do not enjoy the benefit of a carefully constructed set of performance benchmarks''}~\cite{liang2009openrulebench}.

\textbf{CEP and stream processing benchmarks} focus on throughput and latency for data-flow systems.
LinearRoad~\cite{arasu2004linearroad} and NEXMark~\cite{tucker2003nexmark} use synthetic streams with SQL/Beam queries; ESPBench~\cite{hesse2021espbench} extends this to enterprise IoT; Theodolite~\cite{henning2021theodolite} benchmarks cloud-native scalability in Kubernetes.
Among CEP-specific work, Wahl and Hollunder~\cite{wahl2012cep} compare Esper, StreamInsight, and Drools-Fusion on a single SOAP-event pattern; FINCoS~\cite{mendes2013fincos} provides a load-generation framework without a standard workload.
RDF streaming benchmarks (SRBench, CityBench~\cite{ali2015citybench}) incorporate time windows but evaluate SPARQL semantics, not general rules.
None of these provide semantic rule sets, domain-derived event models, or rule-base structural characterisation.

\textbf{The gap.} Authoritative surveys confirm the absence of comprehensive benchmarks at the intersection of rules and streaming.
Cugola and Margara~\cite{cugola2012processing} unify DSMSs and CEP under an Information Flow Processing taxonomy but identify no benchmark exercising both rule semantics and streams.
Dayarathna and Perera~\cite{dayarathna2018event} conclude that \emph{``there are few widely agreed-upon benchmarks for EP; most studies use microbenchmarks.''}
Bizarro et al.~\cite{bizarro2011bicep} explicitly note \emph{``no agreed-upon benchmarks in the CEP community.''}
Luckham~\cite{luckham2002power} defines CEP's conceptual foundations but no standardised evaluation workloads exist.
Our work addresses this gap by combining real-world multi-domain streaming data, rule sets derived from domain standards, full temporal CEP constructs, and a 22-dimension structural characterisation framework grounded in workload analysis principles~\cite{jain1991art}.
Table~\ref{tab:gap} summarises the positioning.

\begin{table}[t]
\centering
\caption{Comparison of existing benchmarks with our proposed suite. (\checkmark = supported, --- = absent)}\label{tab:gap}
\small
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{Benchmark} & \textbf{Rule Type} & \textbf{Data} & \textbf{Temporal} & \textbf{Structure} & \textbf{Year} \\
\midrule
Waltz / Miss Manners    & Production (OPS5)   & Synthetic       & ---             & ---              & 1985 \\
OpenRuleBench           & Datalog/Prod.       & Static KB       & ---             & ---              & 2009 \\
Train Benchmark         & Graph patterns      & Synthetic       & ---             & ---              & 2018 \\
RUBEN                   & Datalog/OWL         & Static KB       & ---             & ---              & 2022 \\
\midrule
LinearRoad / NEXMark    & SQL / Beam          & Synth.\ streams & \checkmark      & ---              & 03--04 \\
Wahl \& Hollunder       & CEP (Esper/Drools)  & Synth.\ SOAP    & \checkmark$^*$  & ---              & 2012 \\
ESPBench / Theodolite   & Beam / Flink        & Synth.\ IoT     & \checkmark      & ---              & 2021 \\
\midrule
\textbf{Our Suite}      & \textbf{Drools CEP} & \textbf{Real-world} & \checkmark  & \checkmark       & 2026 \\
\bottomrule
\multicolumn{6}{@{}l}{\footnotesize $^*$Limited to a single 4-event pattern.}
\end{tabular}
\end{table}
